Yield bias

### 2015-08-27 Rmd/compare-landcover.Rmd
Into results in manuscript. Updated cropland bias maps, bias statistics plots. Added Rmd for miscellaneous supporting calculations.

---
title: "Calculating cropland biases"
author: "Lyndon Estes"
date: "01 October 2014"
output:
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes
---

## Overview
This section of the analysis calculates the differences between the GTI cropland percentages and those from the various landcover products. It draws from results generated by the [cropland pre-processing](https://github.com/PrincetonUniversity/SAcropland/blob/master/vignettes/sa-cropland-analysis.Rmd). (Note to self here: I want to have a transportable link between chapters here, rather than a direct link to the other RMD document in 
github).

## Analyses
### Load in data 
1. Set up paths
```{r, eval = FALSE}
library(lmisc)
library(raster)
library(SAcropland)
p_root <- proj_root("SAcropland")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
```

2. Load cropland grids 
  + Derived from global landcover products (note: it turns out GLC-Share and geowiki are the same thing)

```{r, eval = FALSE}
salc <- raster(full_path(p_data, "sa_ag_masked.tif")) * 100
globsa_list_sum <- lapply(full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
modsa_list_sum <- lapply(full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
glcsa <- raster(full_path(p_data, "glcsa_masked.tif"))
geow <- raster(full_path(p_data, "geowikisa_masked.tif"))
lclist <- c(salc, globsa_list_sum, modsa_list_sum, glcsa, geow)  # into list
names(lclist) <- c("sa30", "globmin", "globmu", "globmax", "modmin", "modmu", "modmax", "glcsh", "geow")
```

  + SA landcover data
```{r, eval = FALSE}
load(full_path(p_data, "MZshapes.Rdata"))  
paths <- full_path(p_data, dir(p_data, paste("cover*.*sum_mask*", sep = "")))
gti <- lapply(paths, raster)
names(gti) <- c("g2007", "g2011")
```
[[back to top]][Overview]

### Calculate differences at different levels of aggregation
This steps draws on functions I wrote to calculate the actual and absolute differences between a given "master" grid and lists of other grids. This comparison is done at multiple aggregations, ranging from 1-600 km resolution cropland averages. The lists have two levels of nesting, here the upper provides the resolution, the inner provides the different landcover datasets.

```{r, eval = FALSE}
# Aggregate rasters
#fact <- c(5, 10, 25, 50, 100, 200, 300)
fact <- 25
lc_agg <- aggregate_rast_list(fact, lclist)   # landcover rasters 
gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters

# Calculate differences
dlist_act <- rast_list_math(list(names(lclist), names(lc_agg), names(gti)),
                            gti_agg, lc_agg, "a - b", silent = FALSE)  # actual
dlist_abs <- rast_list_math(list(names(lclist), names(lc_agg), names(gti)), 
                            gti_agg, lc_agg, "abs(a - b)", silent = FALSE)  # absolute
dlist_gti <- rast_list_math(list("g2007", names(gti_agg), "g2011"), 
                            gti_agg, gti_agg, "a - b", silent = FALSE)

@@

```
[[back to top]][Overview]
### Bias as function of cropcover
+ First create NA mask at 1 km resolution, then weighting raster based on number of non-NA cells in each aggregated pixel, and then "bin" rasters, where each category represents a different cropland fractional cover range (0-5%, 5-10%, etc) 
```{r, eval = FALSE}
# NA mask and weights
namask <- !is.na(gti[[1]])  
area_wgts <- aggregate_rast_list(fact, list(namask), fun = sum)
namask[namask == 0] <- NA
writeRaster(namask, filename = full_path(p_data, "namask.tif"), overwrite = TRUE)

# cropland cover bins
binv <- seq(0, 100, 5)
gti_bins <- lapply(gti_agg, function(x) lapply(x, function(y) cut(y, breaks = binv, include.lowest = TRUE)))
```

+ Calculate biases, pooling MODIS and GlobCover across all version
Step 1. Reshape lists, so that all MODIS and GlobCover variants are pooled by scale
```{r, eval = FALSE}
# Set up indexing parameters
fact_all <- c(1, fact)# full aggregation levels
lev_vec <- paste("f", fact_all, sep = "") 
nm_rt <- c("sa30", "glob", "mod", "glcsh", "geow")

@@

+ Then calculate biases across datasets
```{r, eval = FALSE}
dang <- Sys.time()
dstats_all <- lapply(dlists, function(i) {
  print("processing 1 of 2 main bias lists")
  dstats <- lapply(names(i), function(x) {
    print(paste("..", x))
    l1 <- lapply(lev_vec, function(y) {
      print(paste("....", y))
      d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(j) {
        print(paste("......bin", j))
        d2 <- do.call(rbind, lapply(1:length(i[[x]][[y]]), function(z) {
          print(paste("........variant", z))
          l3 <- i[[x]][[y]][[z]]
          d3 <- do.call(rbind, lapply(1:length(l3), function(k) {
            rs <- gti_bins[[y]][[k]] == j    
            rs[rs == 0] <- NA
            o <- rs * l3[[k]]
            w <- area_wgts[[y]][[1]] * rs
            oo <- cbind.data.frame("lc" = rep(names(i)[z], length(w)), 
                                   getValues(o), getValues(w))
          }))
          d3f <- d3[!is.na(d3[, 2]), ]
        }))
        if(nrow(d2) > (length(i[[x]][[y]]) * 10)) {
        qs <- c(nrow(d2), box_stats(d2[, 2], weighted = TRUE, 
                                    weight.opts = list("weights" = d2[, 3])))
        } else if((nrow(d2) > 0) & (nrow(d2) < length(i[[x]][[y]]) * 20)) {
          qs <- c(rep(NA, 6), Hmisc::wtd.mean(d2[, 2], weights = d2[, 3], 
                                              na.rm = FALSE))
        } else {
          qs <- rep(NA, 7)
        }
      }))
      cbind(binv[-1], d1)
      })
    named_out(l1, lev_vec)
  })
  named_out(dstats, names(i))
})
dut <- Sys.time() - dang  # 33 minutes
save(dstats_all, file = full_path(p_data, "bias_stats.rda"))
```
[[back to top]][Overview]

+ Finally, differences between gti2007 and 2011
```{r, eval = FALSE}
dstats_gti <- lapply(lev_vec, function(x) {
    #x <- lev_vec[[2]]
    i <-  dlist_gti$g2007[[x]][[1]]
    print(paste("..", x))
    d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(y) {
      #y <- 1
      print(paste("....", y))
      rs <- gti_bins[[x]][[2]] == y    
      rs[rs == 0] <- NA
      o <- rs * i
      w <- area_wgts[[x]][[1]] * rs
      oo <- cbind.data.frame(rep(y, length(w)), getValues(o), getValues(w))
      oof <- oo[!is.na(oo[, 2]), ]
      if(nrow(oof) > 10) {
        qs <- c(nrow(oof), box_stats(oof[, 2], weighted = TRUE, 
                                     weight.opts = list("weights" = oof[, 3])))
      } else if((nrow(oof) > 0) & (nrow(oof) < 10)) {
        qs <- c(rep(NA, 6), Hmisc::wtd.mean(oof[, 2], weights = oof[, 3], na.rm = FALSE))
      } else {
        qs <- rep(NA, 7)
      }
    }))
    cbind(binv[-1], d1)
})
save(dstats_gti, file = full_path(p_data, "bias_stats_gti.rda"))

```
Checks: compared number of non-NA cells coming out of each GTI version at each level of aggregation (fine); compared cell counts between previous committed version and this one (difference exists, but due to fact that gtimu is no longer considered; checked that previous versions of difference rasters and reshapes of them were referencing the correct GTI-landcover bias rasters (i.e. to make sure rasters weren't incorrectly indexed: all fine)); box_stats with weighted quantile option compared to previous commit where wtd.quantile was used separately (fine). Compared results from original roughed out version of approach for calculating bias statistics with results from new outputs list (using geowiki at 20 km aggregation as the comparison)

## Next steps
Plot bias results.  
[[back to top]][Overview]

>>>>

---
title: "Calculating cropland biases"
author: "Lyndon Estes"
date: "01 October 2014"
output:
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes
    number_sections: true
---

# Overview
This section of the analysis calculates the differences between the GTI cropland percentages and those from the various landcover products. It draws from results generated by the [cropland pre-processing](cropland-preprocess.Rmd). (Note: html needed).

# Analyses
## Load in data 
### Set up paths
```{r, eval = FALSE}
library(lmisc)
library(raster)
library(SAcropland)
p_root <- proj_root("SAcropland")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
```

### Load cropland grids

Derived from global landcover products (note: it turns out GLC-Share and geowiki are the same thing, so GLC-Share dropped)

```{r, eval = FALSE}
salc <- raster(full_path(p_data, "sa_ag_masked.tif")) * 100
ds <- full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)])
globsa_list_sum <- lapply(ds, raster)
fnames <- full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)])
modsa_list_sum <- lapply(fnames, raster)
# glcsa <- raster(full_path(p_data, "glcsa_masked.tif"))
geow <- raster(full_path(p_data, "geowikisa_masked.tif"))
lclist <- c(salc, globsa_list_sum, modsa_list_sum, geow)  # into list
names(lclist) <- c("sa30", "globmin", "globmu", "globmax", "modmin", "modmu",
                   "modmax", "geow")
```

### SA landcover data
```{r, eval = FALSE}
load(full_path(p_data, "MZshapes.Rdata"))  
paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
gti <- lapply(paths, raster)
names(gti) <- c("g2007", "g2011")
```
<a href="#top">Back to top</a>

## Calculate differences at different levels of aggregation
This steps draws on functions I wrote to calculate the actual and absolute differences between a given "master" grid and lists of other grids. This comparison is done at multiple aggregations, ranging from 1-600 km resolution cropland averages. The lists have two levels of nesting, here the upper provides the resolution, the inner provides the different landcover datasets.

```{r, eval = FALSE}
# Aggregate rasters
fact <- c(5, 10, 25, 50, 100)
#fact <- 25
lc_agg <- aggregate_rast_list(fact, lclist)   # landcover rasters 
gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters

# Calculate differences
# actual
dlist_act <- rast_list_math(list(names(lclist), names(lc_agg), names(gti)),
                            gti_agg, lc_agg, "a - b", silent = FALSE)  
# absolute
dlist_abs <- rast_list_math(list(names(lclist), names(lc_agg), names(gti)), 
                            gti_agg, lc_agg, "abs(a - b)", silent = FALSE)  
dlist_gti <- rast_list_math(list("g2007", names(gti_agg), "g2011"), 
                            gti_agg, gti_agg, "a - b", silent = FALSE)

@@

```
<a href="#top">Back to top</a>

## Bias as function of cropcover
### First create NA mask at 1 km resolution, then weighting raster based on number of non-NA cells in each aggregated pixel, and then "bin" rasters, where each category represents a different cropland fractional cover range (0-5%, 5-10%, etc) 
```{r, eval = FALSE}
# NA mask and weights
namask <- !is.na(gti[[1]])
area_wgts <- aggregate_rast_list(fact, list(namask), fun = sum)
namask[namask == 0] <- NA
writeRaster(namask, filename = full_path(p_data, "namask.tif"), 
            overwrite = TRUE)

# cropland cover bins
binv <- seq(0, 100, 5)
gti_bins <- lapply(gti_agg, function(x) {
  lapply(x, function(y) cut(y, breaks = binv, include.lowest = TRUE))
})
```

### Calculate biases, pooling MODIS and GlobCover across all version
Step 1. Reshape lists, so that all MODIS and GlobCover variants are pooled by scale
```{r, eval = FALSE}
# Set up indexing parameters
fact_all <- c(1, fact)# full aggregation levels
lev_vec <- paste("f", fact_all, sep = "") 
nm_rt <- c("sa30", "glob", "mod", "geow")

@@

### Then calculate biases across datasets

Note: this uses an older and slower version of the code. Subsequent analyses use the faster version based on data.tables.  
```{r, eval = FALSE}
dang <- Sys.time()
dstats_all <- lapply(dlists, function(i) {
  print("processing 1 of 2 main bias lists")
  dstats <- lapply(names(i), function(x) {
    print(paste("..", x))
    l1 <- lapply(lev_vec, function(y) {
      print(paste("....", y))
      d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(j) {
        print(paste("......bin", j))
        d2 <- do.call(rbind, lapply(1:length(i[[x]][[y]]), function(z) {
          print(paste("........variant", z))
          l3 <- i[[x]][[y]][[z]]
          d3 <- do.call(rbind, lapply(1:length(l3), function(k) {
            rs <- gti_bins[[y]][[k]] == j    
            rs[rs == 0] <- NA
            o <- rs * l3[[k]]
            w <- area_wgts[[y]][[1]] * rs
            oo <- cbind.data.frame("lc" = rep(names(i)[z], length(w)), 
                                   getValues(o), getValues(w))
          }))
          d3f <- d3[!is.na(d3[, 2]), ]
        }))
        if(nrow(d2) > (length(i[[x]][[y]]) * 10)) {
        qs <- c(nrow(d2), box_stats(d2[, 2], weighted = TRUE, 
                                    weight.opts = list("weights" = d2[, 3])))
        } else if((nrow(d2) > 0) & (nrow(d2) < length(i[[x]][[y]]) * 20)) {
          qs <- c(rep(NA, 6), Hmisc::wtd.mean(d2[, 2], weights = d2[, 3], 
                                              na.rm = FALSE))
        } else {
          qs <- rep(NA, 7)
        }
      }))
      cbind(binv[-1], d1)
      })
    named_out(l1, lev_vec)
  })
  named_out(dstats, names(i))
})
dut <- Sys.time() - dang  # 33 minutes
save(dstats_all, file = full_path(p_data, "bias_stats.rda"))
```

<a href="#top">Back to top</a>

## Differences between gti2007 and 2011
```{r, eval = FALSE}
dstats_gti <- lapply(lev_vec, function(x) {
    #x <- lev_vec[[2]]
    i <-  dlist_gti$g2007[[x]][[1]]
    print(paste("..", x))
    d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(y) {
      #y <- 1
      print(paste("....", y))
      rs <- gti_bins[[x]][[2]] == y    
      rs[rs == 0] <- NA
      o <- rs * i
      w <- area_wgts[[x]][[1]] * rs
      oo <- cbind.data.frame(rep(y, length(w)), getValues(o), getValues(w))
      oof <- oo[!is.na(oo[, 2]), ]
      if(nrow(oof) > 10) {
        qs <- c(nrow(oof), box_stats(oof[, 2], weighted = TRUE, 
                                     weight.opts = list("weights" = oof[, 3])))
      } else if((nrow(oof) > 0) & (nrow(oof) < 10)) {
        qs <- c(rep(NA, 6), 
                Hmisc::wtd.mean(oof[, 2], weights = oof[, 3], na.rm = FALSE))
      } else {
        qs <- rep(NA, 7)
      }
    }))
    cbind(binv[-1], d1)
})
save(dstats_gti, file = full_path(p_data, "bias_stats_gti.rda"))

```
<a href="#top">Back to top</a>

Checks: compared number of non-NA cells coming out of each GTI version at each level of aggregation (fine); compared cell counts between previous committed version and this one (difference exists, but due to fact that gtimu is no longer considered; checked that previous versions of difference rasters and reshapes of them were referencing the correct GTI-landcover bias rasters (i.e. to make sure rasters weren't incorrectly indexed: all fine)); box_stats with weighted quantile option compared to previous commit where wtd.quantile was used separately (fine). Compared results from original roughed out version of approach for calculating bias statistics with results from new outputs list (using geowiki at 20 km aggregation as the comparison)

### 2015-08-27 Rmd/landcover-bias-plots.Rmd
Into results in manuscript. Updated cropland bias maps, bias statistics plots. Added Rmd for miscellaneous supporting calculations.

## Plotting bias figures
### Biases in space
Load in necessary datasets

@@

Plot maps of biases at different resolutions, for publication output (and display here). To avoid the visual distortion of NA areas filled in by larger levels of aggregation, each aggregated difference must be disaggregated first and then masked before plotting.   
```{r, eval = FALSE}
namask <- raster(full_path(p_data, "namask.tif"))
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
lev <- c("f1", "f25", "f50", "f100")  # just show four levels

@@

Plot of selected rasters and levels 
```{r, eval = FALSE}
# Plotting colors
cols <- colorRampPalette(c("red", "grey80", "blue4"))
brks <- c(-100.1, -75, -50, -25, -10, 0, 10, 25, 50, 75, 100.1)
n_cols <- length(brks) - 1
legtext <- "% Difference"
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev2 <- c("1 km", "25 km", "50 km", "100 km")
pdf(full_path(p_fig, "bias_map.pdf"), height = 6, width = 7)
par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
for(i in 1:length(snms)) {
  print(snms[i])
  for(j in 1:length(lev)) {
    print(lev[j])
    plot(sa.shp, lty = 0)
    plot(disagg[[snms[i]]][[lev[j]]], add = TRUE, col = cols(n_cols), 
         breaks = brks, legend = FALSE)
  if(j == 1) mtext(mcap[i], side = 2, line = 1, cex = cx)
  if(i == 1) mtext(lev2[j], side = 3, line = 0, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext, 
            legend.vals = round(brks), 
            longdims = c(0.2, 0.8), shortdims = c(0.06, 0.01), 
            colvec = cols(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 5), leg.adj = list(c(0.25, 0), c(0.5, -0.5)), 
            cex.val = cx, textcol = lcol, bordercol = lcol)
dev.off()
```
The map below was created by the code above, showing biases at 1 km, 25 km, 50 km, and 100 km from South Africa's landcover product, the mean estimates from 
GlobCover and MODIS, and Geowiki/GLC-SHARE.

@@

[[back to top]][Plotting bias figures]

### Bias in relation to field density
1. Supplementary figures - bias plots at 1 km with boxplots.

```{r, eval = FALSE}
#btype <- c("bias_act.pdf", "bias_abs.pdf")
# plot matrix
nc <- 4
m <- matrix(rep(c(rep(1, 8), rep(2, 8), rep(3, 8), rep(4, 8)), 2), nrow = 6, 
            ncol = 32, byrow = TRUE) 
cols <- c("red", "orange3", "green4", "blue", "yellow")
bcol <- c("grey50", "grey50", "grey50", "grey")
lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
fcol <- "grey"
lev <- names(dstats_all$act$sa30)
lvec <- seq(5, 100, 5)
pp <- list("x" = c(-2, 21), "y" = c(0, 0), 
           "x21" = list(c(0, 9), c(9, 21)),
           "y2" = list(c(seq(-100, -10, 10), seq(10, 100, 10)), 
                       seq(10, 100, 10)), 
           "y1lwd" = list(2, 1),  
           "y2lwd" = 0.5, #c(1, 0.5, 1, 0.5, 0.5, 0.5, 0.5, 1, 0.5, 1),
           "y2lc" = list(c("black", rep("grey80", 4), "black", rep("grey80", 4), 
                           rep("grey80", 4), "black", rep("grey80", 4),"black"),
                         rep("grey80", 10)),
           "yl" = list(c(-100, 100), c(0, 100)), 
           #"yl2" = list(c(-40, 40), c(0, 60)), "xl" = c(1, 20),
           "yl2" = c(-20, 60), "xl" = c(1, 20),
           "xaxl" = unlist(lapply(seq(5, 100, 10), function(x) c(x, " "))),
           "xaxl2" = ifelse(seq(5, 100, 5) %in% c(25, 50, 75), seq(5, 100, 5), 
                            " "),
           "yax1" = list(seq(-100, 100, 50), seq(0, 100, 20)), 
           #"yax2" = list(
            # unlist(lapply(seq(-40, 40, 20), function(x) c(" ", x)))[-1], 
            # unlist(lapply(seq(0, 60, 20), function(x) c(" ", x)))[-1]),
           #"yaxp" = list(-100, 0), 
           "ycol" = c("grey70"))  # plot parameters

@@

2. Bias with scale
```{r, eval=FALSE}
lev <- lev[!lev %in% c("f5", "f300")]
levnms <- paste(gsub("f", "", lev), "km")
m2 <- unlist(lapply(seq(1, by = 1, length = length(lev) * 4), function(x) {
  rep(x, nc)
}))
m3 <- do.call(rbind, lapply(seq(1, length(m2), length(lev) * nc), function(x) {
  do.call(rbind, rep(list(m2[x:(x + length(lev) * nc - 1)]), nc))
}))

@@

pdf(full_path(p_fig, "biases_1-200km.pdf"), width = 7.5, height = 7)
l <- layout(m3)
# i <- 1; j <- 1
par(mar = c(0.5, 0, 1, 0.3), oma = c(5, 4, 2, 2))
for(i in 1:length(snms)) {
  highcounter <- rep(0, length(lev))
  for(j in 1:length(lev)) {
    plot(pp$x, pp$y, ylim = pp$yl2, xlim = pp$xl, type = "l", lwd = 1, 
         axes = FALSE, xlab = "", ylab = "", xaxs = "i")
    for(z in 1:length(yl)) {
      lines(pp$x, rep(yl[z], 2), lwd = pp$y2lwd, col = ylcol[z])
    }
    muv <- rep(NA, 2)
    for(bt in 1:2) {
      d <- dstats_all[[bt]][[i]][[j]]
      n <- nrow(d)
      lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = ylwd[bt], 
            lty = lt[bt])
      muv[bt] <- round(mean(d[, 8], na.rm = TRUE))
      toohigh <- which(d[, "mu"] > pp$yl2[2])
      if(length(toohigh) > 0) {
        highcounter[j] <- 1
        xi <- toohigh[which.max(d[toohigh, "mu"])]
        text(xi, y = pp$yl2[2] + 4, labels = round(d[xi, "mu"]), cex = 1.1,
             col = "grey50", xpd = NA)
      }
    }
    text(5, pp$yl2[2] - 14, labels = paste(muv, collapse = " / "), cex = 1.1, 
         col = "grey10")
    if(j == 1) axis(2, at = yl, labels = yl, las = 2, mgp = c(1, 0.6, 0))
    if(j == 1) mtext(lcnms[i], side = 2, line = 2.8, cex = 1, col = cols[i])
    if(i == 1) mtext(levnms[j], side = 3, line = -0.4, cex = 1.2)
    axis(1, at = c(1, 5, 10, 15, 20), labels = rep("", 5), mgp = c(1, 1, 0), 
         tcl = -0.7, lwd.ticks = 1.2, pos = pp$yl2[1])
    axis(1, at = 1:20, labels = rep("", 20), las = 2, mgp = c(1, 0.4, 0), 
         tcl = -0.3, pos = pp$yl2[1])
    if(i == length(snms)) {
      axis(1, at = 1:20, labels = pp$xaxl2, las = 2, mgp = c(1, 0.8, 0), 
           tcl = -0.3, pos = pp$yl2[1])
    }
  }
}
mtext("% bias", side = 2, line = 2.1, cex = 0.8, outer = TRUE, adj = 0.5)
mtext("% cropland", side = 1, outer = TRUE, line = 1.5, cex = 0.8)
par(xpd = NA)
x <- grconvertX(0.75, from = "ndc", to = "user")
y <- grconvertY(0.07, from = "ndc", to = "user")
legend(x = x, y = y, legend = c("actual bias", "absolute bias"), lty = lt, 
       lwd = ylwd, bty = "n", cex = 1.2)
dev.off()

DF <- do.call(rbind.data.frame, lapply(snms[4], function(x) {
  dstats_all[[2]][[x]]$f1[, c(1, 8)]
}))
dlm <- lm(mu ~ poly(V1, degree = 2), data = DF)
plot(DF[, 1], DF[, 2])
lines(predict.lm(dlm, newdata = data.frame("V1" = 1:110)))
```

>>>>

# Plotting bias figures
# Spatial biases
## Datasets

@@

## Plots
Plot maps of biases at different resolutions, for publication output (and display here). To avoid the visual distortion of NA areas filled in by larger levels of aggregation, each aggregated difference must be disaggregated first and then masked before plotting.   
```{r, eval = FALSE}
namask <- raster(full_path(p_data, "namask.tif"))
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
lev <- c("f1", "f25", "f50", "f100")  # just show four levels

@@

## Plot of selected rasters and levels 
```{r, eval = FALSE}
# Plotting colors
cols <- colorRampPalette(c("red", "grey80", "blue4"))
brks <- c(-100.1, -75, -50, -25, -10, -5, -1, 1, 5, 10, 25, 50, 75, 100.1)
brklen <- length(brks) - 1
cols <- cols(brklen)
cols[c(1, length(cols))] <- c("darkred", "purple3")
legtext <- "% Difference"
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev2 <- c("1 km", "25 km", "50 km", "100 km")
pdf(full_path(p_fig, "bias_map.pdf"), height = 7, width = 7)
par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(3.7, 1.5, 1.5, 0))
for(i in 1:length(snms)) {
  print(snms[i])
  for(j in 1:length(lev)) {
    print(lev[j])
    plot(sa.shp, lty = 0)
    plot(disagg[[snms[i]]][[lev[j]]], add = TRUE, col = cols, breaks = brks,
         legend = FALSE)
  if(j == 1) mtext(mcap[i], side = 2, line = -0.25, cex = cx)
  if(i == 1) mtext(lev2[j], side = 3, line = -0.5, cex = cx)
  }
}
flex_legend(ncuts = brklen, legend.text = legtext, 
            legend.vals = round(brks), longdims = c(0.2, 0.8), 
            shortdims = c(0.05, 0.01), colvec = cols, srt = c(270, 0), 
            horiz = TRUE, textside = "bottom", legend.pos = c(4, 5), 
            leg.adj = list(c(0.2, 0.35), c(0.5, -0.5)), cex.val = cx, 
            textcol = lcol, bordercol = lcol)
dev.off()
```
The map below was created by the code above, showing biases at 1 km, 25 km, 50 km, and 100 km from South Africa's landcover product, the mean estimates from 
GlobCover and MODIS, and Geowiki/GLC-SHARE.

@@

<a href="#top">Back to top</a>

# Supplementary figures: bias in relation to field density
## Bias plots at 1 km with boxplots.

```{r, eval = FALSE}
#btype <- c("bias_act.pdf", "bias_abs.pdf")
# plot matrix
nc <- 4
m <- matrix(rep(c(rep(1, 8), rep(2, 8), rep(3, 8), rep(4, 8)), 2), nrow = 6, 
            ncol = 32, byrow = TRUE) 
cols <- c("red", "orange3", "green4", "blue", "yellow")
bcol <- c("grey50", "grey50", "grey50", "grey")
lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
fcol <- "grey"
lev <- names(dstats_all$act$sa30)
lvec <- seq(5, 100, 5)
pp <- list("x" = c(-2, 21), "y" = c(0, 0), 
           "x21" = list(c(0, 9), c(9, 21)),
           "y2" = list(c(seq(-100, -10, 10), seq(10, 100, 10)), 
                       seq(10, 100, 10)), 
           "y1lwd" = list(2, 1),  
           "y2lwd" = 0.5, #c(1, 0.5, 1, 0.5, 0.5, 0.5, 0.5, 1, 0.5, 1),
           "y2lc" = list(c("black", rep("grey80", 4), "black", rep("grey80", 4), 
                           rep("grey80", 4), "black", rep("grey80", 4),"black"),
                         rep("grey80", 10)),
           "yl" = list(c(-100, 100), c(0, 100)), 
           "yl2" = c(-20, 60), "xl" = c(1, 20),
           "xaxl" = unlist(lapply(seq(5, 100, 10), function(x) c(x, " "))),
           "xaxl2" = ifelse(seq(5, 100, 5) %in% c(25, 50, 75), seq(5, 100, 5), 
                            " "),
           "yax1" = list(seq(-100, 100, 50), seq(0, 100, 20)), 
           "ycol" = c("grey70"))  # plot parameters

@@

## Bias with scale
```{r, eval=FALSE}
#lev <- lev[!lev %in% c("f5", "f300")]
levnms <- paste(gsub("f", "", lev), "km")
m2 <- unlist(lapply(seq(1, by = 1, length = length(lev) * 4), function(x) {
  rep(x, nc)
}))
m3 <- do.call(rbind, lapply(seq(1, length(m2), length(lev) * nc), function(x) {
  do.call(rbind, rep(list(m2[x:(x + length(lev) * nc - 1)]), nc))
}))

@@

pdf(full_path(p_fig, "biases_1-100km.pdf"), width = 7.5, height = 7)
l <- layout(m3)
# layout.show(l)
# i <- 1; j <- 1
par(mar = c(0.5, 0, 1, 0.3), oma = c(5, 4, 2, 2))
for(i in 1:length(snms)) {
  highcounter <- rep(0, length(lev))
  for(j in 1:length(lev)) {
    plot(pp$x, pp$y, ylim = pp$yl2, xlim = pp$xl, type = "l", lwd = 1, 
         axes = FALSE, xlab = "", ylab = "", xaxs = "i")
    for(z in 1:length(yl)) {
      lines(pp$x, rep(yl[z], 2), lwd = pp$y2lwd, col = ylcol[z])
    }
    muv <- rep(NA, 2)
    for(bt in 1:2) {
      d <- dstats_all[[bt]][[i]][[j]]
      n <- nrow(d)
      lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = ylwd[bt], 
            lty = lt[bt])
      muv[bt] <- round(mean(d[, 8], na.rm = TRUE))
      toohigh <- which(d[, "mu"] > pp$yl2[2])
      if(length(toohigh) > 0) {
        highcounter[j] <- 1
        xi <- toohigh[which.max(d[toohigh, "mu"])]
        text(xi, y = pp$yl2[2] + 4, labels = round(d[xi, "mu"]), cex = 1.1,
             col = "grey50", xpd = NA)
      }
    }
    text(5, pp$yl2[2] - 14, labels = paste(muv, collapse = " / "), cex = 1.1, 
         col = "grey10")
    if(j == 1) axis(2, at = yl, labels = yl, las = 2, mgp = c(1, 0.6, 0))
    if(j == 1) mtext(lcnms[i], side = 2, line = 2.8, cex = 1, col = cols[i])
    if(i == 1) mtext(levnms[j], side = 3, line = -0.4, cex = 1.2)
    axis(1, at = c(1, 5, 10, 15, 20), labels = rep("", 5), mgp = c(1, 1, 0), 
         tcl = -0.7, lwd.ticks = 1.2, pos = pp$yl2[1])
    axis(1, at = 1:20, labels = rep("", 20), las = 2, mgp = c(1, 0.4, 0), 
         tcl = -0.3, pos = pp$yl2[1])
    if(i == length(snms)) {
      axis(1, at = 1:20, labels = pp$xaxl2, las = 2, mgp = c(1, 0.8, 0), 
           tcl = -0.3, pos = pp$yl2[1])
    }
  }
}
mtext("% bias", side = 2, line = 2.1, cex = 0.8, outer = TRUE, adj = 0.5)
mtext("% cropland", side = 1, outer = TRUE, line = 1.5, cex = 0.8)
par(xpd = NA)
x <- grconvertX(0.75, from = "ndc", to = "user")
y <- grconvertY(0.07, from = "ndc", to = "user")
legend(x = x, y = y, legend = c("actual bias", "absolute bias"), lty = lt, 
       lwd = ylwd, bty = "n", cex = 1.2)
dev.off()

```

```{r, eval=FALSE, echo=FALSE}
DF <- do.call(rbind.data.frame, lapply(snms[4], function(x) {
  dstats_all[[2]][[x]]$f1[, c(1, 8)]
}))
dlm <- lm(mu ~ poly(V1, degree = 2), data = DF)
plot(DF[, 1], DF[, 2])
lines(predict.lm(dlm, newdata = data.frame("V1" = 1:110)))
```

### 2015-08-27 Rmd/yield-bias.Rmd
Into results in manuscript. Updated cropland bias maps, bias statistics plots. Added Rmd for miscellaneous supporting calculations.

# Assume a yield gap over-estimated by 100% in 400 mm isohyet
ygap_bias2 <- sapply(lev, function(x) {
  if(x != "f1") { 
    preagg <- aggregate(prebin, fact = as.numeric(gsub("f", "", x)), fun = modal)
  } else {
    preagg <- prebin
  }
  prez <- preagg == 400
  notprez <- preagg != 400
  sapply(snms, function(y) {
    yg <- (g1 * cyld_agg[[x]]$gti * notprez) + (g2 * cyld_agg[[x]]$gti * prez)
    yg2 <- (g1 * cyld_agg[[x]][[y]] * notprez) + (g2 * cyld_agg[[x]][[y]]*prez)
    pg <- tareas[[x]] * carea_agg[[x]]$gti * yg
    pg2 <- tareas[[x]] * carea_agg[[x]][[y]] * yg2
    cellStats(pg - pg2, sum) / cellStats(pg, sum) * 100
  })
})

>>>>

# Assume a yield gap over-estimated by 100% in 400 mm isohyet
ygap_bias2 <- sapply(lev, function(x) {
  if(x != "f1") { 
    preagg <- aggregate(prebin, fact = as.numeric(gsub("f", "", x)), fun=modal)
  } else {
    preagg <- prebin
  }
  prez <- preagg == 400
  notprez <- preagg != 400
  sapply(snms, function(y) {
    yg <- (g1 * cyld_agg[[x]]$gti * notprez) + (g2 * cyld_agg[[x]]$gti * prez)
    yg2 <- (g1 * cyld_agg[[x]][[y]] * notprez) + (g2 * cyld_agg[[x]][[y]]*prez)
    pg <- tareas[[x]] * carea_agg[[x]]$gti * yg
    pg2 <- tareas[[x]] * carea_agg[[x]][[y]] * yg2
    cellStats(pg - pg2, sum) / cellStats(pg, sum) * 100
  })
})
